Experience Report: Distributed Computing for the Common Man

### 2010-03-29 doc/icfp-2010/main.tex
*** empty log message ***

\usepackage{amsmath,graphicx, color}

@@

Our library, \functory, 
In this section, we show the potential of our library using two 
computationally-extensive examples.

@@

The first column shows the value of $N$.
Each computation is divided into a certain number of tasks,
which is shown in second column.
Then the last three columns show execution timings in seconds for the
three implementations. The figures within brackets show the speedup
ratio w.r.t. sequential implementation.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    N & \#tasks  & sequential& cores                 & network 
    \\\hline\hline
    16 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
    \\\hline
       &  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
    \\\hline
    17 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
    \\\hline
       &  240    & 107.0     &  14.00 (7.64$\times$) & 24.90  (4.30$\times$)
    \\\hline
    18 &   18    & 787.0     & 123.00 (6.40$\times$) & 125.00 (6.30$\times$)  
    \\\hline
       &  272    & 787.0     & 103.00 (7.64$\times$) & 124.00 (6.34$\times$)  
    \\\hline
    19 &   19    &6120.0     & 937.00 (6.53$\times$) & 940.00 (6.51$\times$)  
    \\\hline
       &  306    &6130.0     & 796.00 (7.70$\times$) & 819.00 (7.48$\times$)
    \\\hline
  \end{tabular}
\end{center}
From the table above, it is clear that the \of{Cores} and \of{Network}
implementations provide a significant speedup. As evident from the
last row, the speedup ratio is almost 8, which is also the number of
cores we use.  It is also evident from the last column that
\of{Network} implementation performs significantly better when the
computation time dominates the total execution time.  The two extreme
cases correspond to second and last rows: in the second row, the
communication time dominates and is in fact more than 91\%\ of the
total execution time; on the other hand, for the last row
communication time amounts to just 4.6\%\ of the total execution time.
As expected, the network implementation is only beneficial when the
computation time for each individual task is significant, which is the
case in realistic examples.

@@

The master program proceeds by reading the file names, turning them
into tasks by multiplying them by the number of provers (more than
300 tasks in total).
Each worker in turn invokes the given prover on the given file.
We set a timeout limit of 1 minute.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{cannot-decide} (depending on
whether the VC is valid or undecided by the prover), 
\emph{timeout} and \emph{failure}.
The figure below shows the total time in minutes spent by each prover
for each possible outcomes.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    prover   & valid & cannot & timeout & failure
    \\\hline\hline
    Alt-ergo &  24.7  &  2.4  & 1260.0 &   0.0       
    \\\hline
    Simplify &   0.6  &  0.3  &  300.0 &  43.0   
    \\\hline
    Z3       &  98.6  &  0.0  & 1200.0 &  66.3    
    \\\hline
    CVC3     &  46.7  & 28.0  &  720.0 &   0.0   
    \\\hline
  \end{tabular}
\end{center}
These figures sum up to more than 1 hour if provers were executed
sequentially. However, using our library and our 3-machine
infrastructure, it completes in 4 minutes and 8 seconds, giving us a
speedup of more than 15$\times$.

@@

% \acks

% Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

>>>>

\usepackage{amsmath,graphicx,color,url}

@@

Our library, \functory, is implemented in \Ocaml\ and is available
from \url{http://www.lri.fr/~filliatr/functory/}.
In this section, we show the potential of this library using two 
computationally-extensive examples.

@@

The first column shows the value of $N$.  Each computation is divided
into a certain number of tasks, by placing some of the queens before
starting the distributed computation.  The number of tasks is shown in
second column.  Then the last three columns show execution timings in
seconds for the three implementations. The figures within brackets
show the speedup ratio w.r.t. sequential implementation.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    N & \#tasks  & sequential& cores                 & network 
    \\\hline\hline
    16 &   16    &  15.2     &   2.04 (7.45$\times$) &  2.35  (6.47$\times$) 
    \\\hline
       &  210    &  15.2     &   2.01 (7.56$\times$) & 21.80  (0.69$\times$)
    \\\hline
    17 &   17    & 107.0     &  17.20 (6.22$\times$) & 16.20  (6.60$\times$)
    \\\hline
       &  240    & 107.0     &  14.00 (7.64$\times$) & 24.90  (4.30$\times$)
    \\\hline
    18 &   18    & 787.0     & 123.00 (6.40$\times$) & 125.00 (6.30$\times$)  
    \\\hline
       &  272    & 787.0     & 103.00 (7.64$\times$) & 124.00 (6.34$\times$)  
    \\\hline
    19 &   19    &6120.0     & 937.00 (6.53$\times$) & 940.00 (6.51$\times$)  
    \\\hline
       &  306    &6130.0     & 796.00 (7.70$\times$) & 819.00 (7.48$\times$)
    \\\hline
  \end{tabular}
\end{center}
From the table above, it is clear that the \of{Cores} and \of{Network}
implementations provide a significant speedup. As evident from the
last row, the speedup ratio is almost 8, which is also the number of
cores we use.  It is also evident from the last column that
\of{Network} implementation performs significantly better when the
computation time dominates the total execution time.  The two extreme
cases correspond to second and last rows: in the second row, the
communication time dominates and is in fact more than 91\%\ of the
total execution time; on the other hand, for the last row
communication time amounts to just 4.6\%\ of the total execution time.
As expected, the network implementation is only beneficial when the
computation time for each individual task is significant, which is the
case in realistic examples.

@@

The master program proceeds by reading the file names, turning them
into tasks by multiplying them by the number of provers (more than
300 tasks in total).
Each worker in turn invokes the given prover on the given file.
We set a timeout limit of 1 minute.
Each task completes with one of the four possible outcomes: \emph{valid},
\emph{cannot-decide} (depending on
whether the VC is valid or undecided by the prover), 
\emph{timeout} and \emph{failure}.
The figure below shows the total time in minutes spent by each prover
for each possible outcomes.
\begin{center}
  \begin{tabular}{|r|r|r|r|r|}
    \hline
    prover   & valid & cannot & timeout & failure
    \\\hline\hline
    Alt-ergo &  24.7  &  2.4  & 1260.0 &   0.0       
    \\\hline
    Simplify &   0.6  &  0.3  &  300.0 &  43.0   
    \\\hline
    Z3       &  98.6  &  0.0  & 1200.0 &  66.3    
    \\\hline
    CVC3     &  46.7  & 28.0  &  720.0 &   0.0   
    \\\hline
  \end{tabular}
\end{center}
These figures sum up to more than 1 hour if provers were executed
sequentially. However, using our library and our 3-machine
infrastructure, it completes in 4 minutes and 8 seconds, giving us a
speedup of more than 15$\times$. We are still far away from the ideal
ratio of 20$\times$ (we are using 20 cores), since some provers are
allocating a lot of memory and time spent in system calls is not
accounted for in the total observed time. However, a ratio of
15$\times$ is already a significant improvement for our day-to-day
experiments.

@@

\acks
The authors are grateful to the ProVal team for support and comments
on early versions of the library and of this paper.

\vfill\pagebreak
% We recommend abbrvnat bibliography style.
